---
title: "Physical Activity and Weight in Highschoolers"
author: "Kushal Kundanmal"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(plotly)

```
# Physical Activity and Weight in Highschool students.
In this analysis, we are going to study whether physical activity per week is related to weight in a sample of High School Students.

## The data

Every two years, the Centers for Disease Control and Prevention conduct the [Youth Risk Behavior Surveillance System (YRBSS)](https://www.cdc.gov/healthyyouth/data/yrbs/index.htm) survey, where it takes data from high schoolers (9th through 12th grade), to analyze health patterns. We will work with a selected group of variables from a random sample of observations during one of the years the YRBSS was conducted.

## Load the data

We first load the data that we will work with. This data is part of the `openintro` textbook. There are observations on 13 different variables, some categorical and some numerical. To get started, we first use `glimpse()` to gain initial insights into the data set.


```{r}
data(yrbss)
glimpse(yrbss)
```


## Exploratory Data Analysis

In a similar fashion, it is useful to check with `skim()` to get a feel for missing values, summary statistics of numerical variables, and a very rough histogram.

Looking at the weights variable, we see that we are missing 1004 observations. The distribution is right skewed, so with most observations clustering to the left.

```{r, eda_on_weight}
skim(yrbss)
```

Next, we consider the possible relationship between a high schooler’s weight and their physical activity. Plotting the data is a useful first step because it helps us quickly visualize trends, identify strong associations, and develop research questions.

We create a new variable `physical_3plus`, which will be `yes` if they are physically active for at least 3 days a week, and `no` otherwise.

```{r, physical_variable}
# Create new variable physical_3plus, which indicates if a high schooler is physically active for at least 3 days a week or not

yrbss <- yrbss %>% 
  mutate(physical_3plus = ifelse(physically_active_7d >= 3, "yes", "no"))

yrbss_summary <- yrbss %>% 
  filter(!is.na(physical_3plus)) %>% 
  group_by(physical_3plus) %>% 
  summarise(count = n()) %>% 
  mutate(prop= count/sum(count))
yrbss_summary
```

Diving deeper into `physical_3plus`, we provide a 95% confidence interval for the high schoolers that are not active for at least 3 days a week.


```{r, confidence_interval}

# Sum population
count_pop <- sum(yrbss_summary$count)

# Calculating the SE
proportion_sample <- yrbss_summary$prop[1]
se_pop <- sqrt((proportion_sample * (1 - proportion_sample))/count_pop)

# Calculating the mean
mean_pop <- se_pop/sqrt(count_pop)

# t-critical value for 5% significance
t_critical <- qt(0.975, count_pop - 1)
lower_ci <- mean_pop - t_critical * se_pop
upper_ci <- mean_pop + t_critical * se_pop
print(paste("The 95% confidence interval for the population proportion of high schoolers that are not active 3 or more days per week is: [", round(lower_ci, digits=3), ",", round(upper_ci, digits=3), "].", sep = ""))

```

We expected physical activity to be quite influential on the weight of high schoolers. However, when looking at the box plot, the difference in weight for high schoolers who exercise more than 3 hours a week and for those who exercise less is not profound. Even more surprising is the observation that those that exercise more than 3 hours actually have a slightly higher weight. In addition, the two groups have roughly the same amount and size of outliers in weight values.

```{r, boxplot}
#boxplot
yrbss_clean <- na.omit(yrbss)
ggplot(data = yrbss_clean, aes(x = physical_3plus, y= weight)) +
  geom_boxplot() +
  labs(title="Enough weekly exercise does not affect a high schooler's weight", 
       subtitle="Relationship between weight and weekly physical exercise for high schoolers", 
       x="Weekly physical exercise of 3 hours or more", 
       y="Weight in kilograms")
 
#calculating the mean and the median of the weight
yrbss_clean %>%  group_by(physical_3plus)%>% summarize(mean  = mean(weight), median = median(weight))
```

## Confidence Interval

Boxplots show how the medians of the two distributions compare, but we can also compare the means of the distributions using either a confidence interval or a hypothesis test. Below, we display the mean, sd, se, count, t-critical value, margin of error, lower confidence bound, and upper confidence bound for the variable `weight` - where we distinguish again between those high schoolers who are active for at least 3 days a week and those who are not.


```{r}
yrbss %>%
  group_by(physical_3plus) %>%
  filter(!is.na(physical_3plus)) %>% 
  summarise(mean_weight = mean(weight, na.rm = TRUE),
            sd_weight = sd(weight, na.rm=TRUE),
            count = n(),
            se_weight = sd_weight/sqrt(count),
            t_critical = qt(0.975, count-1), 
            margin_of_error = t_critical * se_weight,
            lower = mean_weight - t_critical * se_weight,
            upper = mean_weight + t_critical * se_weight
            )

```

There is an observed difference of about 1.77kg (68.44 - 66.67), and we notice that the two confidence intervals do not overlap. It seems that the difference is at least 95% statistically significant. Let us also conduct a hypothesis test.

## Hypothesis test with formula
The null hypothesis in this case states that the mean of the first group is equal to the mean of the second group. In other words, the difference in means equals zero. The alternative hypothesis states that these means are not equal.

```{r}
t.test(weight ~ physical_3plus, data = yrbss)
```

## Hypothesis test with `infer`
Next, we will introduce a new function, `hypothesize`, that falls into the infer workflow. We will use this method for conducting hypothesis tests. First, we need to initialize the test, which we will save as `obs_diff`.

```{r}
obs_diff <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

```

Now, we simulate the test on the null distribution, which we will save as null.

```{r}
null_dist <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

```


Here, `hypothesize` is used to set the null hypothesis as a test for independence, i.e., that there is no difference between the two population means. Also, the `type` argument within generate is set to permute, which is the argument when generating a null distribution for a hypothesis test. We can visualize this null distribution with the following code:

```{r}
ggplot(data = null_dist, aes(x = stat)) +
  geom_histogram()

```

Finally, we visualise the test and look at the `obs_diff`.
We see that obs_diff = 1.77

```{r}

null_dist %>% visualize() +
  shade_p_value(obs_stat = obs_diff, direction = "two-sided")

null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")

```
## Conclusion
We can see that there is a statistically significant difference in means between the two groups. On average, the group that exercises 3 or more times per week is between 2.42kg to 1.12kg heavier than the group that does not exercise at least 3 times per week.
